---
title: "Red Wine Exploration"
author: "Olivia Jonah"
date: "March 9, 2017"
output: html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```

```{r }
#setwd("C:/Users/t609110/Documents/OJ/Udacity")
#getwd()
```
###Introduction

This study is going to analyze wine data created by Paulo Cortez (Univ. Minho),
Antonio Cerdeira, Fernando Almeida, Telmo Matos and Jose Reis (CVRVV) @ 2009.
it shows how certain varibles can affect the quality of wine.

```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(dplyr)
library(memisc)
library(DAAG)
library(Simpsons)
library(pander)
library(lazyeval)
theme_set(theme_minimal(20))
```

```{r  Load_the_Data}
wine <- read.csv('wineQualityReds.csv')
#wine$rating <- NULL

# Transform quality from an integer to a ordered factor
wine$quality <- factor(wine$quality, ordered = T)

# create a rating varible
wine$rating <- ifelse(wine$quality < 5, 'bad', ifelse(
  wine$quality < 7, 'average', 'good'))
wine$rating <- ordered(wine$rating,
                     levels = c('bad', 'average', 'good'))
```

# Univariate Plots Section
##Database Structure
The dataset contains objetive and subjective quality data for 1599 red wines. 
There are a total of 14 variables, of which 13 are objective quality factors 
obtained from quality tests such as pH test, 
and 1 subjective factor that contains median expert evaluation score

```{r  results='asis', Variables}
print("Dataset variables")
pandoc.list(names(wine))
```


```{r  Structure}
str(wine)
print("Dataset structure")
```



```{r  results='asis', Summary}
set.caption("Dataset summary")
pandoc.table(summary(wine))
```

To be able to choose what assumptions can be made 
when applying statistical tests, 
each individual distribution will be explored.

```{r  Rating}
ggplot(data=wine, aes(x=quality)) +
   stat_count(binwidth = 1)
ggplot(data=wine, aes(x=rating)) +
   stat_count()
```

The plot above shows a high average ratings of wines
(high concentration of wines in the center region)
and there are noticabley no outliers, 
this might be a problem for generating a predicting model later on.

```{r  fixed.acidity}
ggplot(data=wine, aes(x=fixed.acidity)) +
  geom_histogram(binwidth = 1)
```

There is a high concentration of wines with fixed.acidity close to 8 
(themedian) mean up to 9.2

```{r  volatile.acidity}
qplot(volatile.acidity, data = wine, binwidth = .01)
```

This shows a normal distribution

```{r  citric.acid}
qplot(citric.acid, data = wine, binwidth = .01)
```

Now this is strange distribution.
8% of wines do not present critic acid at all.
Maybe a problem in the data collection process?


```{r  residual.sugar}
qplot(residual.sugar, data = wine, binwidth = .1)
```

Median 2.2 ( high concentration) with some outliers along
the higher ranges.


```{r  chlorides}
qplot(chlorides, data = wine, binwidth = .01)
```

We see a similar distribution with chlorides.
where in about 8% has no chlorides

```{r  free.sulfur.dioxide}
qplot(free.sulfur.dioxide, data = wine, binwidth = 1)
```

The distributions peaks at around aroun 7 
and from then on resembles a long tailed
distribution with very few wines over 60.

```{r  total.sulfur.dioxide}
qplot(total.sulfur.dioxide, data = wine, binwidth = 1)
```

long tailed distribution
```{r warning=FALSE, density}
qplot(density, data = wine, binwidth = 0.001)
```

This is a normal distribution for density.

```{r  pH}
qplot(pH, data = wine, binwidth = 0.01) +
  scale_x_continuous(breaks=c(2.8,3,3.2,3.4,3.6,3.8,4))
```

pH also looks normally distributed.

```{r  sulphates}
qplot(sulphates, data = wine, binwidth = .1)
```

For sulphates we see a distribution similar to the ones of residual.sugar and
chlorides.

```{r  warning=FALSE, alcohol}
qplot(alcohol, data = wine, binwidth = .1)
```

Rapid increase and then long tailed distribution as seen in
sulfur.dioxide, maybe there is a correlation between the variables.

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 observation of wines in the dataset with 14 features . There is
one categorical variable (quality) and the others are numerical variables 
indicating  physical and chemical properties of the wine.

Other observations:
The median quality is 6, which in the given scale (1-10) is a mediocre wine.
The better wine in the sample has a score of 8, and the worst has a score of 3.
The dataset is not balanced, that is, there are a more average wines than poor
or excelent ones and this might prove challenging when designing a predicting 
algorithm.

### What is/are the main feature(s) of interest in your dataset?
The main feature in the data is quality. I'd like to determine which features
determine the quality of wines.

### What other features in the dataset do you think will help support your 
investigation into your feature(s) of interest?
The variables related to acidity (fixed, volatile, citric.acid and pH) might
explain some of the variance. Maybe the different acid concentrations might
alter the taste of the wine.
Also, residual.sugar dictates how sweet a wine is and might also have an
influence in taste.

### Did you create any new variables from existing variables in the dataset?
A rating variable was created to improve the later visualizations.

### Of the features you investigated, were there any unusual distributions? Did 
you perform any operations on the data to tidy, adjust, or change the form of
the data? If so, why did you do this?

Citric.acid stood out from the other distributions. It had (apart from some
outliers) an retangularly looking distribution which given the wine quality
distribution seems very unexpected.

# Bivariate Plots Section

A correlation table for all variables will help us understand the relationships
between them.

```{r results='asis', Correlation_Matrix}
# in order to calculate the correlation matrix we need to transform the data
# a little
c <- cor(
  wine %>%
    # first we remove unwanted columns
    dplyr::select(-rating) %>%
    mutate(
      # now we translate quality to a number
      quality = as.numeric(quality)
    )
  )

emphasize.strong.cells(which(abs(c) > .3 & c != 1, arr.ind = TRUE))
pandoc.table(c)
```
 
Alcohol has negative correlation with density as alcohol is
less dense than water.

Volatile.acidity has a positive correlation with pH as pH
is a direct measure of acidity.
Maybe the effect of a lurking variable( outliers)

Residual.sugar does not show correlation with quality.
Free.sulfur.dioxide and total.sulfur.dioxide are highly correlated as expected.

Density has a very strong correlation with fixed.acidity.
The variables that have the strongest correlations to quality are
volatile.acidity and alcohol.

Using boxplots to further examine the relationship between some varibles
and quality.

```{r echo=FALSE}
# generate a function to output summaries along with the charts
summarise_boxplot <- function(data, x, y) {
  pandoc.table(data %>%
  group_by_(x) %>%
  summarise_(
    mean = interp(~mean(v), v=as.name(y)),
    median = interp(~median(v), v=as.name(y))
    ), caption = paste("Summaries for", y, "grouped by", x))
}
```

```{r echo=FALSE, Quality_vs_fixed.acidity}
ggplot(data=wine, aes(x=quality, y=fixed.acidity)) +
  geom_boxplot()
```

```{r  results='asis', Quality_vs_fixed.acidity_summaries}
summarise_boxplot(wine, "quality", "fixed.acidity")
```

As the correlation table showed, fixed.acidity seems to have little to no effect
on quality.

```{r echo=FALSE, Quality_vs_volatile.acidity}
ggplot(data=wine, aes(y=volatile.acidity, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_volatile.acidity_summaries}
summarise_boxplot(wine, "quality", "volatile.acidity")
```

volatile.acidity has an inverse relationship.
Quality seems to go up when volatile.acidity goes down.
The higher ranges seem to produce more average and poor wines.

```{r echo=FALSE, Quality_vs_citric.acid}
ggplot(data=wine, aes(x=quality, y=citric.acid)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_citric.acid_summaries}
summarise_boxplot(wine, "quality", "citric.acid")
```

A direct proportionality relationship between these two variables.
Better wines tend to have higher concentration of citric acid.

```{r echo=FALSE, Quality_vs_residual.sugar}
ggplot(data=wine, aes(x=as.factor(quality),
                      y=residual.sugar)) +
  geom_boxplot() + 
  coord_cartesian(ylim=c(0, 30))
```

```{r echo=FALSE, results='asis', Quality_vs_residual.sugar_summaries}
summarise_boxplot(wine, "quality", "residual.sugar")
```

Residual.sugar  seems to have little to no effect 
on perceived quality.

```{r echo=FALSE, Quality_vs_chlorides}
ggplot(data=wine, aes(y=chlorides, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_chlorides_summaries}
summarise_boxplot(wine, "quality", "chlorides")
```

A lower concentration of chlorides seem to produce
better wines.

```{r echo=FALSE, Quality_vs_free.sulfur.dioxide}
ggplot(data=wine, aes(y=free.sulfur.dioxide, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_free.sulfur.dioxide_summaries}
summarise_boxplot(wine, "quality", "free.sulfur.dioxide")
```

Though the  ranges are really close to each other 
it seems that too little sulfur dioxide equals a  
poor quality of wine and too much gives an average quality wine. 

```{r echo=FALSE, Quality_vs_total.sulfur.dioxide}
ggplot(data=wine, aes(y=total.sulfur.dioxide, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_total.sulfur.dioxide_summaries}
summarise_boxplot(wine, "quality", "total.sulfur.dioxide")
```

```{r echo=FALSE, free_vs_total.sulfur.dioxide}
ggplot(data=wine, aes(y=total.sulfur.dioxide, x=free.sulfur.dioxide)) +
  geom_point(alpha = 0.8, size = 2) +
      geom_smooth(method = "lm", se = FALSE,size=1)
```

here is a very similar distribution as this is a set
of free.sulfur.dioxide.

```{r echo=FALSE, Quality_vs_density}
ggplot(data=wine, aes(y=density, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_density_summaries}
summarise_boxplot(wine, "quality", "density")
```

For better wines quality the density should be lower
this is probably due to the alcohol concentration. 
Maybe we might want to see the effect id alcohol is constant.

```{r echo=FALSE, Quality_vs_pH}
ggplot(data=wine, aes(y=pH, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_pH_summaries}
summarise_boxplot(wine, "quality", "pH")
```

Better quality wines ten to be more acidic

Let's examine how each acid concentration affects pH.

```{r echo=FALSE}
ggplot(data = wine, aes(x = fixed.acidity, y = pH)) +
  geom_point() +
  geom_smooth(method="lm")
ggplot(data = wine, aes(x = fixed.acidity, y = pH)) +
  geom_point() +
  scale_x_log10(breaks=seq(5,15,1)) +
  xlab("log10(fixed.acidity)") +
  geom_smooth(method="lm")
ggplot(data = wine, aes(x = volatile.acidity, y = pH)) +
  geom_point() +
  geom_smooth(method="lm")
ggplot(data = wine, aes(x = volatile.acidity, y = pH)) +
  geom_point() +
  scale_x_log10(breaks=seq(.1,1,.1)) +
  xlab("log10(volatile.acidity)") +
  geom_smooth(method="lm")
ggplot(data = wine, aes(x = citric.acid, y = pH)) +
  geom_point() +
  geom_smooth(method="lm")
ggplot(data = subset(wine, citric.acid > 0), aes(x = citric.acid, y = pH)) +
  geom_point() +
  scale_x_log10() +
  xlab("log10(citric.acid)") +
  geom_smooth(method="lm")
```

Hmm surprisingly the acid concentration would has a positive
correlation with pH. Maybe Simpsons Paradox?

```{r include=FALSE, Simpsons}
simpsons <- Simpsons(volatile.acidity, pH, data=wine)
```

```{r echo=FALSE, Simpsons.plot}
plot(simpsons)
```

to know if Simpsons Paradox is present, the regression coeficient
is recalculated and the data is clusterized. the change in sign
indicates that there is a lurking variable distorting the overall coefficient.

Since ph measures acid concentration using log scale,
therefore the correlation between ph and the log of acid concentration
is stronger.
A linear model can help us tell the variance in ph level of the acidity variables

```{r}
str(df)
```


```{r echo=FALSE, warning=FALSE, Linear_Model_Acids}
acidsLM <- lm(pH ~ 
                I(log10(citric.acid)) +
                I(log10(volatile.acidity)) +
                I(log10(fixed.acidity)),
                    data = subset(wine,citric.acid > 0))
summary(acidsLM)
df = data.frame(wine$quality, wine$pH )
df$predictions <- predict(acidsLM, wine)
df$error <- (df$predictions - wine$pH)/wine$pH

ggplot(data=df, aes(x=wine.pH, y=error)) +
  geom_boxplot()

```

this shows that the medians (which generally will be close to the average) are all at the same level. 

```{r echo=FALSE, Quality_vs_sulphates}
ggplot(data=wine, aes(y=sulphates, x=quality)) +
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_sulphates_summaries}
summarise_boxplot(wine, "quality", "sulphates")
```

Good quality wines seems to have higher sulphates 
concentration.

```{r echo=FALSE, Quality_vs_alcohol}
ggplot(data=wine, aes(y=alcohol, x=quality)) + 
  geom_boxplot()
```

```{r echo=FALSE, results='asis', Quality_vs_alcohol_summaries}
summarise_boxplot(wine, "quality", "alcohol")
```

Here we see an increase in alcohol graduation gives an
increase in the concentration of better graded wines.
Given the high number of outliers it seems we cannot rely on alcohol alone to
produce better wines.

Let's try using a simple linear model to investigate.

```{r echo=FALSE, Linear_Model_over_alcohol}
alcoholQualityLM <- lm(as.numeric(quality) ~ alcohol,
                    data = wine)
summary(alcoholQualityLM)
df = data.frame(wine$quality)
df$predictions <- predict(alcoholQualityLM, wine)
df$error <- (df$predictions - as.numeric(wine$quality))/as.numeric(wine$quality)

ggplot(data=df, aes(x=wine.quality, y=error)) +
  geom_boxplot()
```

It seems  seems alcohol alone only explains about 22% of
the variance in quality. Other variables needs to be explored to
generate a better model.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the
investigation. How did the feature(s) of interest vary with other features in
the dataset?
Here are my observations:

Fixed.acidity seems to have little to no effect on quality

Volatile acidity has and inverse relation

Citirc acidity ith higer concentration produces 
better quality in wines

it was surprising to see that residula sugar has little
or no effect on the quality od wine


Eventtough the corelation is not too strong, 
a lower concentration of chlorides seem to produce
better wines.

Better wines tend to have lower densities.

pH:it seems better wines are more acid (but there were
many outliers). Better wines also seem to have a
higher concentration of sulphates.

the Linear model tells that if some other variables 
were added it might generate a bette rmodel,
but we can substantiate that Alcohol graduation has a 
strong correlation with quality, but it cannot explain 
all the variance.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

there is a strong relation between free and total sulfur.dioxide.

Linear model showed volitalie acidity has a positive coefficient 
for the relation between the acid concentration and pH.

### What was the strongest relationship you found?

The relationship between the variables total.sulfur.dioxide and
free.sulfur.dioxide.

# Multivariate Plots Section

### Alcohol and other variables

Using multivariate plots to answer some questions encountered
and to look for other relationships in the data.

```{r echo=FALSE, warning=FALSE, density_over_alcohol_vs_quality}
ggplot(aes(x = alcohol, 
           y = sulphates  , color = factor(quality)), 
       data = wine) +
      geom_point(alpha = 0.8, size = 2) +
      geom_smooth(method = "lm", se = FALSE,size=1, color="darkblue")  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality'))

```

alcohol is kept constant and it shows that density does not affect 
quality of wines


```{r echo=FALSE, warning=FALSE, sulphates_vs_alcohol_vs_quality}
ggplot(data = wine,
       aes(y = sulphates, x = alcohol,
           color = quality)) +
  geom_point() +
  scale_y_continuous(limits=c(0.3,1.5)) +
  scale_color_brewer()
ggplot(data = wine,
       aes(y = sulphates, x = alcohol,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE,size=1, color="darkblue")  +
  scale_y_continuous(limits=c(0.3,1.5)) +
  facet_wrap(~rating) +
  scale_color_brewer()
```

worth noting thatbfor wines with high alcohol content, 
having a higher concentration of sulphates will produce
better wines.

```{r echo=FALSE, warning=FALSE, volatile.acidity_vs_alcohol_vs_quality}
ggplot(data = wine,
       aes(y = volatile.acidity, x = alcohol,
           color = quality)) +
  geom_point() +
  scale_color_brewer()
ggplot(data = wine,
       aes(y = volatile.acidity, x = alcohol,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE,size=1,
              color="darkblue")  +
  facet_wrap(~rating) +
  scale_color_brewer()
```

For volatile acidity,having less acetic acid on
higher concentration of alcohol seems to produce better wines.

```{r echo=FALSE, warning=FALSE, pH_vs_alcohol_vs_quality}
ggplot(data = wine,
       aes(y = pH, x = alcohol,
           color = quality)) +
  geom_point() +
  scale_color_brewer()
ggplot(data = wine,
       aes(y = pH, x = alcohol,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE,size=1,
              color="darkred")  +
  facet_wrap(~rating) +
  scale_color_brewer()
```

Low pH and high alcohol concentration seem to work well.

### Acid exploration

Using multivariate plots to further investigate the
relationship between the acids and quality.

```{r echo=FALSE, fixed.acidity_vs_volatile.acidity_vs_quality}
ggplot(data = wine,
       aes(y = fixed.acidity, x = volatile.acidity,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2,) +
  geom_smooth(method = "lm", se = FALSE,size=1,
              color="darkred")  +
  scale_color_brewer()
```

Almost no variance in the y axis compared to the x axis. Lets try the other
acids.

```{r echo=FALSE, citric.acid_vs_volatile.acidity_vs_quality}
ggplot(data = wine,
       aes(y = citric.acid, x = volatile.acidity,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE,size=1,
              color="darkred")  +
  facet_wrap(~rating) +
  scale_color_brewer()
```

High citric acid and low acetic acid seems like a good combination.

```{r echo=FALSE, citric.acid_vs_fixed.acidity_vs_quality}
ggplot(data = wine,
       aes(y = citric.acid, x = fixed.acidity,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE,size=1,
              color="darkred")  +
  scale_color_brewer()
  cor.test(wine$citric.acid, wine$fixed.acidity) 
```

Although there seems to be a correlation between tartaric acid and citric acid
concentrations, nothing stands out in terms of quality.
```{r echo=FALSE, Linear_Model}
set.seed(87654321)
training_data <- sample_frac(wine, .6)
test_data <- wine[ !wine$X %in% training_data$X, ]
m1 <- lm(as.numeric(quality) ~ alcohol, data = training_data)
m2 <- update(m1, ~ . + sulphates)
m3 <- update(m2, ~ . + volatile.acidity)
m4 <- update(m3, ~ . + citric.acid)
m5 <- update(m4, ~ . + fixed.acidity)
m6 <- update(m2, ~ . + pH)
mtable(m1,m2,m3,m4,m5,m6)
df <- data.frame(
  test_data$quality,
  predict(m5, test_data) - as.numeric(test_data$quality)
)
names(df) <- c("quality", "error")
ggplot(data=df, aes(x=quality,y=error)) +
  geom_boxplot(col="darkgreen") 
   
```




# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

if one wants a very good wine, one should have a high alcohol content
and high sulphate concentrations combined 

### Were there any interesting or surprising interactions between features?

yes alcohol and density. one does not affect the other.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

- variables: alcohol, sulphates, and the acid variables.
Both the residuals plot and the cross validation favors average wines.
This is probably a reflection of the high number of average wines in the
training dataset or it could mean that there is missing information that would
help predict the edge cases.

Maybe the dataset has some errors

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, warning=FALSE, alcohol.final}
ggplot(data=wine, aes(y=alcohol, x=quality)) + 
  geom_boxplot() +
  xlab("Quality ") +
  ggtitle("Influence of alcohol on wine quality")
```

### Description One
This boxplot has different wine quality levels on its x-axis to the amount of alcohol on its y-axis. 
It clearly shows the increasing trend in the amount of alcohol when wine quality moves from lower level to higher level. 
The plot validates the concept of "balancing taste" attribute of good quality wine
Box plots allow for comparison of data from different categories for easier,
more effective decision-making.


### Plot Two

```{r echo=FALSE, warning=FALSE, sulphates_vs_alcohol_vs_quality.final}
ggplot(data = wine,
       aes(y = sulphates, x = alcohol,
           color = quality)) +
  geom_point(alpha = 0.8, size = 2) +
  geom_smooth(method = "lm", se = FALSE,size=1,
              color="darkred")  +
  scale_y_continuous(limits=c(0.3,1.5)) +
  ylab("potassium sulphate (g/dm3)") +
  xlab("alcohol (% by volume)") +
  scale_color_brewer() +
  ggtitle("Alcohol and sulphates over wine quality")
```

### Description Two
Using a  scatterplot, this graph is aimed at identifying the unique "content profile" of quality wines
by combining the two most important factors (sulphates and alcohol). 
 The graph shows that as wine quality improves, with increase in alcohol and sulphate.
To be able to  ignore the outliers  just eyeball the cloud of dots to locate a general trend,
a scatter plot with regression line was used which helps me to make a discernable connection 
between the 3 things investigate dhere (alcohol, sulphates and quality)
Combining the two High alcohol contents and high sulphate concentrations seem to produce better wines.

### Plot Three

```{r echo=FALSE, final_linear_model_residuals}
df <- data.frame(
  test_data$quality,
  predict(m5, test_data) - as.numeric(test_data$quality)
)
names(df) <- c("quality", "error")
ggplot(data=df, aes(x=quality,y=error)) +
  geom_boxplot(col="darkgreen") +
  ggtitle("Box plot model errors over expected quality")
```


### Description Three

Similiar to plot one, this plot uses a boxplot to show the different attributes across different quality levels.
However, a minor improvement is made by changing the y-axis to show the error. 
Though small, the change leads to a much more clear comparison amongst all the levels.
This change is driven by the idea that the higher quality has minimun error



------
# Reflection

The wine data set contains information on the chemical properties of a selection
of wines collected in 2009. It also includes sensorial data (wine ranking).

By looking at the individual distributions of the variables,
to see how they relate to quality.

it was quick to  notice the high concentration of wines in the middle
ranges of the ranking, that is, average tasting wines.this raise eyebrows
of a true correlation between the variables or maybe for lack of outlier.
None of the correlations were above 0.7, however.
The two variables that had relatively strong correlations
were alcohol percentage and volatile acidity, 
but the individual correlations were not strong 
enough to make definitive conclusions with only bivariate analysis methods
Citric acid stood out because as research shows citric acid is sometimes
added to wines to boost overall acidity.

[researching the subject](https://en.wikipedia.org/wiki/Acids_in_wine#Citric_acid)

having said that the plots shows a strange distribution with some peaks but
showing an almost rectangular distribution specially in the 0-0.5 range.

All of the other variables showed either an normal or long tailed looking
distribution.

Next step was to see how the variables relate with each other to produce 
poor or better quality


This shows alcohol concentration, sulphates and the
individual acid concentrations.

I also tried investigating the effect of each acid in the overall pH for the
wine. I used scatterplots to explore the relationships graphically and also
generated a linear model to check how much of pH the three variables accounted
for.

Having a positive ph for correlation between  acetic
acid concentration suggests the presence of some lurking variable
(Simpsons paradox) using"Simpsons" package I confirmed that suspicion.

Finding out that the concentration of the three acids
only account for less than half of the variance in pH, tells that
there more components affecting acidity that were not measured.

Finally multivariate plots were used to find out interesting combinations 
of variables that might affect quality, and to confirm that density did not
have an effect on quality whilst alcohol concentration constant.

the increase in quality has lesser error.
The graphs adequately illustrate the factors that make good wines 'good' and bad wines 'bad'

For future studies, including malic and lactic acid would be interesting 
to mesure as these improves the taste.
This data set may include storage time and storage method since these factors
can influence the quality of wine as well.
Also, maybe the relationship between price and quality of wine 
to investigate whether expensive wines lead to better quality.
[Wikipedia](https://en.wikipedia.org/wiki/Acids_in_wine) for example,

Maybe the dataset can have entry for the individuals judgemnt
on the flavor, aroma (if can be clasiified), to 
make an interesting analysis.

### Reference
1.Udacity teaching  material
2.research on Simpsons paradox
3.discussion forum on udacity
4.stackoverflow
5.google
6.wikipedia
